---
title: "github-intro-exercise"
author: "Eliana Paradise"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(palmerpenguins)
library(tidyverse)
library(rmarkdown)
library(ggthemes)
library(ggdark)
library(viridis)
library(plotly)
library(RColorBrewer)
```

```{r}
packageLoad <-
  function(x) {
    for (i in 1:length(x)) {
      if (!x[i] %in% installed.packages()) {
        install.packages(x[i])
      }
      library(x[i], character.only = TRUE)
    }
  }

packages <- c('tidyverse',
              'palmerpenguins',
              'lterdatasampler')

packageLoad(packages)
```


```{r}
data("knz_bison")
head(knz_bison)
```

Create a new column in the bison dataset that calculates the individual’s age at the time of observation.
```{r}
# Convert obs_date to Date format if needed
knz_bison <- knz_bison %>%
  mutate(rec_year = year(rec_year),
         age_at_obs = animal_yob - rec_year)

head(knz_bison[, c("animal_yob", "rec_year")])

```


Write a function that will first filter the bison dataset based on any user-specified weight threshold (returning the filtered dataset).
```{r}
filter_by_weight <- function(data, threshold) {
  filtered <- data %>% filter(animal_weight >= threshold)
  return(filtered)
}

# Example filtering "heavy bison" by excluding those who weigh less than 500 kg
heavy_bison <- filter_by_weight(knz_bison, 500)
head(heavy_bison)

```


Create a boxplot showing weight distribution between male and female bison, and then conduct a t-test to see if weight is significantly difference between males and females.

Because of the very small p-value, the weight distribution between males and females is statistically significant. 

```{r}
# Boxplot
ggplot(knz_bison, aes(x = animal_sex, y = animal_weight, fill = animal_sex)) +
  geom_boxplot() +
  labs(title = "Weight Distribution by Sex", x = "Sex", y = "Weight (kg)") +
  theme_minimal()

# T-test
t_test_result <- t.test(animal_weight ~ animal_sex, data = knz_bison)
t_test_result

```


